---
title: "Weight Lifting Exercises Analysis"
author: "Daniel A Stearns"
date: "May 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
##install.packages("doParallel")

echo = TRUE
options(scipen = 1)  
knitr::opts_chunk$set(cache=TRUE)
set.seed(256)

## load libraries
library("caret")
library("randomForest")
library("Hmisc")
library("foreach")
library("doParallel")

## Clear all environment variables for a clean workspace

## Set Working Directory
setwd("~/Coursera/practical machine learning/PML_Assignment1")

```
#Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

#Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

#Processing
Load the data sets for analysis.
```{r}
src_training <- read.csv("pml-training.csv", header = TRUE, na.strings = c("NA", "", "#DIV/0!"));
src_testing <- read.csv("pml-testing.csv", header = TRUE, na.strings = c("NA", "", "#DIV/0!"));
```

#Cleansing
Clean the data set to remove unnecessary columns and columns with bad data.
```{r}
valid_attributes <- colnames(src_training[colSums(is.na(src_training)) == 0])[-(1:7)];
training_data <- src_training[valid_attributes];
testing_data <- src_testing[valid_attributes[valid_attributes!='classe']];
```

#Analysis
Review the data set to see what we are working with.
```{r}
dim(training_data);
colnames(training_data);
head(training_data);

```

#Partioning
Partioning the training data set into two data sets, 60% for training data set  and 40% for testing data set:.
```{r}
partition <- createDataPartition(y=training_data$classe, p=0.6, list=FALSE )
training_set <- training_data[partition,]
testing_set <- training_data[-partition,]
```

#Learning
Use Random Forests for ML.
```{r}
rf_ml <- randomForest(training_set[-ncol(training_set)], training_set$classe)

predictions_training <- predict(rf_ml, newdata=training_set)
confusionMatrix(predictions_training,training_set$classe)

predictions_testing <- predict(rf_ml, newdata=testing_set)
confusionMatrix(predictions_testing,testing_set$classe)

predictions <- predict(rf_ml, newdata=testing_data)
```

#Conclusions
it can be determined that the Randon Forest Machine Learning Algorithm yields an accuracy of 99.51% in determining the exercise is correctly performed.

#Appendix
Generating Testing Files 
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(predictions)
```